Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration

### 2011-11-29 paper/self_calibration_I.tex
euclid: paper after David comments
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19957 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\shorttitle{Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
\shortauthors{Holmes et al.}

@@

\author{R. Holmes\altaffilmark{1} and H.-W. Rix\altaffilmark{1}}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg, 69117, Germany.}

\and

\author{D. W. Hogg\altaffilmark{2}}
\affil{2 Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}

\begin{abstract}
In this paper we show that, with due care given to the observing strategy, it will be possible to accurately constrain the relative photometric calibration of instruments used in large imaging surveys with their science data alone. We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky according to a defined survey strategy. These catalog based simulations include realistic measurement uncertainties and a complex, position dependent instrument response. We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions. By considering four simple survey strategies we find that, with a correct redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of the resulting dataset. The majority of the remaining post self-calibration errors are due to the limitations in the basis used to model the relative instrument response. We find that returning the same sources to very different focal plane positions is the key property of a survey strategy that is required for an accurate calibration. From the results of this work, we are able to highlight an important point for those considering the design of large scale imaging surveys: depart from a regular tiling of the sky and return the same sources to very different focal plane positions.
\end{abstract}

@@

%% From the front matter, we move on to the body of the paper.
%% In the first two sections, notice the use of the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below. We have
%% chosen the first three characters of the first author's name plus
%% the last two numeral of the year of publication as our KEY for
%% each reference.

%% Authors who wish to have the most important objects in their paper
%% linked in the electronic edition to a data center may do so by tagging
%% their objects with \objectname{} or \object{}.  Each macro takes the
%% object name as its required argument. The optional, square-bracket 
%% argument should be used in cases where the data center identification
%% differs from what is to be printed in the paper.  The text appearing 
%% in curly braces is what will appear in print in the published paper. 
%% If the object name is recognized by the data centers, it will be linked
%% in the electronic edition to the object data available at the data centers  
%%
%% Note that for sources with brackets in their names, e.g. [WEG2004] 14h-090,
%% the brackets must be escaped with backslashes when used in the first
%% square-bracket argument, for instance, \object[\[WEG2004\] 14h-090]{90}).
%%  Otherwise, LaTeX will issue an error.

\section{Introduction}
Astronomers tend to think in terms of taking science data and calibration data separately. The former is used for science and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on. But typically far more photons are collected during science exposures themselves; are these not incredibly constraining on the calibration? Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}. But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.

In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies optimized from the very start with this kind of ``self-calibration'' in mind. This work focuses on the relative photometric calibration of a typical imaging instrument only, although we also suspect that similar techniques could be used to constrain many other calibration parameters. The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield the same flux values. Through end-to-end, catalog level survey simulations\footnote{All of the code used in this work is publicly available at XX.}, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.

We construct realistic survey datasets through mock observations of a synthetic sky according to a specified survey strategy. These simulations include a complex, position dependent instrument response for the imaging instrument. Through the self-calibration procedure, we then recover this instrument response by fitting a model that best describes the survey dataset. By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies. There is a degeneracy in this problem: the self-calibration procedure is only able to fit for a \textit{relative} instrument response. We have no way to know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.

In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey datasets. Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure. In Section \ref{sec:simple_surveys} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties. We do not produced pixelated images; instead we concentrate on catalog level simulations with realistic measurement uncertainties. Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.

\section{Survey Dataset Simulations}
\label{sec:survey_simulations}
We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy. In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated. The simulations are split into a number of steps. The first of which is the generation of a synthetic sky. With a given pointing, single mock observations are performed on this sky. To build up a survey wide dataset, multiple single observations are performed according to the specified survey strategy. In this section, we detail the assumptions and methods used in each of these steps.

\subsection{The Synthetic Sky}
We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and $10\upsigma{}$ limits of deep, space-based, near-infrared survey. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
\begin{eqnarray}
\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
\end{eqnarray}
where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters. Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citep{win11} only. These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$. We intentionally omit the stellar population, so that the conclusions drawn from these simulations are independent of the final position of the survey area on the sky. The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy). Due to computational reasons, we only select the brightest sources within the survey area, up to a source density $d$, for the self-calibration procedure.

@@

\subsubsection{Measured Count Rates}
We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ 
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response, and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

\subsubsection{Noise Model}
\label{sec:noise}
To construct the noise model, the simulated exposures are assumed to be background limited and that, for systematic reasons, there is an upper limit on the signal-to-noise ratio of 500 for bright sources. The noise model is complicated further by applying an extra term $\epsilon_i$ to the count rates' uncertainty variance, which we intentionally do not take into account in the analysis in order to simulate systematic problems with the instrument noise model. The \textit{true} noise model is therefore
\begin{eqnarray}
\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
\end{eqnarray}
where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$. The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are assumed to be
\begin{eqnarray*}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2} \, c^{2}_i \quad 
\end{eqnarray*}
during this stage.

@@

\subsection{The Complete Survey}
In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting dataset. We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy. The resultant source measurements are collaged into a survey wide dataset.

\section{Self-Calibrating the Survey Wide Dataset}
\label{sec:self_cal}
We self-calibrate the dataset generated in the survey simulations to recover the true instrument response applied and the true source fluxes. This self-calibration procedure has been successfully applied to ground based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''}. This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates. These steps are iterated until the system converges, or until it is clear that the system will not converge. There is a degeneracy in the problem, as both the true instrument response and the true source magnitudes are unknown. It is therefore only possible to calibrate the \textit{relative} instrument response and the \textit{relative} source fluxes. It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower. In practice, this degeneracy can be broken through observations of standard absolute sources.

@@

@@

\subsection{Step 2: Instrument Response Refinement}
The instrument response parameters can now be refined with the latest source flux estimates. A error function for all the measurements of all the sources is constructed
\begin{eqnarray*}
\chi^2 = \sum_{k} \chi'^2_k \quad ,
\end{eqnarray*}
where
\begin{eqnarray*}
\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i}^2}  \quad .
\end{eqnarray*}
Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eight order polynomial. This can be expressed as
\begin{eqnarray*}
f_{i}(\vec{x_i} | \vec{q}) = \sum_{l = 1}^L q_{l} \, g_l(\vec{x_i})  \quad ,
\end{eqnarray*}
where $L = 45$ in this case. The total error function $\chi^2$ can be rewritten as
\begin{eqnarray*}
\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i}^2}   \quad .
\end{eqnarray*}
To refine the instrument response model fit, this error function is minimized with respect to the instrument response model parameters $q_{l}$
\begin{eqnarray*}
\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0  \quad ,
\end{eqnarray*}
\begin{eqnarray*}
\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i}^2}   \quad .
\end{eqnarray*}
It is now simpler to proceed in matrix notation. The following substitutions can be made
\begin{eqnarray}
b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i^2}    \quad ,\\
G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i^2} g_l(x_i) g_{l'}(x_i)   \quad .
\end{eqnarray}
The matrix equation is then
\begin{eqnarray*}
\vec{b} = G \cdot \vec{q'}    \quad .
\end{eqnarray*}
The refined instrument response parameters are then found by
\begin{eqnarray*}
\vec{q'} \leftarrow G^{-1}  \cdot \vec{b}   \quad .
\end{eqnarray*}
These two steps are iterated until the solution converges to a final fit of the instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ and the source fluxes $s_{k,\fit}$, or until it is clear that a solution will not be found.

\section{Metrics}
\label{sec:metrics}
To assess the performance of the self-calibration procedure with different survey strategies, it is necessary to quantify the quality of the final fitted solution. To do this we defined three quantities. The first is the root-mean-squared (RMS) error $S_\text{RMS}$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for the $K$ sources used in the self-calibration procedure:
\begin{equation}
S_\text{RMS} = \sqrt{\dfrac{1}{K} \sum_k^K \left( \dfrac{s_{k,\fit} - s_{k,\true}}{s_{k,\true}} \right)^2}   \quad .
\end{equation}
The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $300 \times 300$ grid across the focal plane. For the ``True Badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points  
\begin{equation}
B_\true = \sqrt{\dfrac{1}{J} \sum_j^J \left(\dfrac{f_\fit(\vec{x_j} | \vec{q_\fit}) - f_\true(\vec{x_j} | \vec{q_\true})}{f_\true(\vec{x_j} | \vec{q_\true})}\right)^2}   \quad .
\end{equation}
The ``Best-in-Basis Badness'' $B_\basis$ compares the \textit{fitted} instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ to the \textit{best instrument response fit possible} $f_\basis(\vec{x_j} | \vec{q_\basis})$ with the basis used to describe the fitted model (in this case an eight order polynomial) at the $J$ sample points
\begin{equation}
B_\basis = \sqrt{\dfrac{1}{J} \sum_j^J \left( \dfrac{f_\fit(\vec{x_j} | \vec{q_\fit}) - f_\basis(\vec{x_j} | \vec{q_\basis})}{f_\basis(\vec{x_j} | \vec{q_\basis})} \right)^2}   \quad .
\end{equation}
The badnesses provide a more complete description of the self-calibration performance than the RMS error on the fitted sources' fluxes, as the RMS source error only applies to the bright sources within the survey selected for the self-calibration procedure.

@@

\subsection{Self-Calibration Performance}
The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets. With the Survey A dataset it does not converge, even when the system is started close to the optimum fit. The fitted instrument response solutions from the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}. In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``Best-in-Basis''). The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.

@@

With the Survey C and D datasets, the instrument response is accurately recovered. As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$. The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.

>>>>

\shorttitle{Designing Large Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
\shortauthors{Holmes et al.}

@@

\author{Rory Holmes, David W. Hogg\altaffilmark{1} and Hans-Walter Rix}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg, 69117, Germany.}
\altaffiltext{1}{2 Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}

\begin{abstract}
In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate instruments used in large-scale imaging surveys with survey science data alone. We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky according to a defined survey strategy. These catalog based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response. We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions. By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed. The majority of the remaining post-self-calibration errors are due to the limitations in the basis used to model the relative instrument response. We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate calibration. The results of this work suggest the following advice for those considering the design of large scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
\end{abstract}

@@

\section{Introduction}
Astronomers tend to think in terms of taking science data and calibration data separately. The former is used for scientific measurements of astronomical sources and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on. But typically far more photons, or readout electrons, are collected during science exposures themselves; are these not incredibly constraining on the calibration? Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}. But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy is required in the data stream, and a redundancy of a very specific kind.

In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies designed from the very start with this kind of ``self-calibration'' in mind. This work focuses on the relative photometric calibration of a typical imaging instrument only, although similar techniques could be used to constrain many other calibration parameters, such as the optical distortion, point-spread-functions and so-on. The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values. Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.

We construct realistic survey datasets through mock observations of a synthetic sky according to a specified survey strategy. These simulations include a complex, position-dependent instrument response for the imaging instrument. Through the self-calibration procedure, we recover this instrument response by fitting a model that best describes the survey dataset. By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies. There is a degeneracy in this problem: the self-calibration procedure is only able to fit for a \textit{relative} instrument response. We have no way to know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.

In this work we do not produced pixelated images; instead we concentrate on catalog-level simulations with realistic measurement uncertainties. Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.

In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey datasets. Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure, with Section \ref{sec:metrics} summarizing the metrics we use to assess its performance. In Section \ref{sec:simple_surveys} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.

\section{Survey Dataset Simulations}
\label{sec:survey_simulations}
We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy. In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated. The simulations are split into a number of steps. The first is the generation of a synthetic sky. With a given pointing, single mock observations are performed on this sky. To build up a full-survey dataset, multiple single observations are performed according to the specified survey strategy. In this section, we detail the assumptions and methods used in each of these steps.

\subsection{The Synthetic Sky}
We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and $10\upsigma{}$ limits of deep, space-based, near-infrared survey. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
\begin{eqnarray}
\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
\end{eqnarray}
where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters. Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citet{win11} only. These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$. The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy). To limit the data, we only select the brightest sources within the survey area, up to a source density per unit solid angle $d$, for the self-calibration procedure.

@@

\subsubsection{Measured Count Rates}
We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ 
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response, and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

\subsubsection{Noise Model}
\label{sec:noise}
To construct the noise model, the simulated exposures are assumed to be background limited and that, for systematic reasons, there is an upper limit on the signal-to-noise ratio of 500 for bright sources. The noise model is complicated further by applying an extra term $\epsilon_i$ to the count rates' uncertainty variance, which we intentionally do not take into account in the analysis in order to simulate systematic problems with the instrument noise model. The \textit{true} noise model is therefore
\begin{eqnarray}
\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
\end{eqnarray}
where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$. The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are \textit{assumed} (incorrectly, but realistically) to be
\begin{eqnarray*}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2} \, c^{2}_i \quad 
\end{eqnarray*}
during this stage.

@@

\subsection{The Complete Survey}
In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting dataset. We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy. The resultant source measurements are collated into a survey-wide dataset.

\section{Self-Calibrating the Survey-Wide Dataset}
\label{sec:self_cal}
We self-calibrate the dataset generated in the survey simulations to recover a best possible estimate of the true instrument response applied and the true source fluxes. This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''}. This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates. These steps are iterated until the system converges, or until it is clear that the system will not converge. There is a degeneracy in the problem, as both the true instrument response and the true source magnitudes are unknown. It is therefore only possible to calibrate the \textit{relative} instrument response and the \textit{relative} source fluxes. It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower. In practice, this degeneracy can be broken through observations of a small number of standard absolute sources.

@@

We make one further assumption during this analysis phase: We assume that clearly variable sources have been removed from the source catalog, and that the remaining sources vary by less than $\epsilon_{max}$.

@@

\subsection{Step 2: Instrument Response Refinement}
The instrument response parameters can now be refined with the latest source flux estimates. A error function for all the measurements of all the sources is constructed
\begin{eqnarray*}
\chi^2 = \sum_{k} \chi'^2_k \quad ,
\end{eqnarray*}
where
\begin{eqnarray*}
\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i}^2}  \quad .
\end{eqnarray*}
Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eighth order polynomial. This can be expressed as
\begin{eqnarray*}
f_{i}(\vec{x_i} | \vec{q}) = \sum_{l = 1}^L q_{l} \, g_l(\vec{x_i})  \quad ,
\end{eqnarray*}
where $L = 45$ in this case. The total error function $\chi^2$ can be rewritten as
\begin{eqnarray*}
\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i}^2}   \quad .
\end{eqnarray*}
To refine the instrument response model fit, this error function is minimized with respect to the instrument response model parameters $q_{l}$
\begin{eqnarray*}
\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0  \quad ,
\end{eqnarray*}
\begin{eqnarray*}
\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i}^2}   \quad .
\end{eqnarray*}
It is now simpler to proceed in matrix notation. The following substitutions can be made
\begin{eqnarray}
b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i^2}    \quad ,\\
G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i^2} g_l(x_i) g_{l'}(x_i)   \quad .
\end{eqnarray}
The matrix equation is then
\begin{eqnarray*}
\vec{b} = G \cdot \vec{q'}    \quad .
\end{eqnarray*}
The refined instrument response parameters are then found by
\begin{eqnarray*}
\vec{q'} \leftarrow G^{-1}  \cdot \vec{b}   \quad .
\end{eqnarray*}
These two steps are iterated until the solution converges to a final fit of the instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ and the source fluxes $s_{k,\fit}$, or until it is clear that a solution will not be found.

\section{Metrics}
\label{sec:metrics}
To assess the performance of the self-calibration procedure with different survey strategies, it is necessary to quantify the quality of the final fitted solution. To do this we defined three quantities. The first is the root-mean-squared (RMS) error $S_\text{RMS}$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for the $K$ sources used in the self-calibration procedure:
\begin{equation}
S_\text{RMS} = \sqrt{\dfrac{1}{K} \sum_k^K \left( \dfrac{s_{k,\fit} - s_{k,\true}}{s_{k,\true}} \right)^2}   \quad .
\end{equation}
The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $300 \times 300$ grid across the focal plane. For the ``true badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points  
\begin{equation}
B_\true = \sqrt{\dfrac{1}{J} \sum_j^J \left(\dfrac{f_\fit(\vec{x_j} | \vec{q_\fit}) - f_\true(\vec{x_j} | \vec{q_\true})}{f_\true(\vec{x_j} | \vec{q_\true})}\right)^2}   \quad .
\end{equation}
The ``best-in-basis badness'' $B_\basis$ compares the \textit{fitted} instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ to the \textit{best instrument response fit possible} $f_\basis(\vec{x_j} | \vec{q_\basis})$ with the basis used to describe the fitted model (in this case an eight order polynomial) at the $J$ sample points
\begin{equation}
B_\basis = \sqrt{\dfrac{1}{J} \sum_j^J \left( \dfrac{f_\fit(\vec{x_j} | \vec{q_\fit}) - f_\basis(\vec{x_j} | \vec{q_\basis})}{f_\basis(\vec{x_j} | \vec{q_\basis})} \right)^2}   \quad .
\end{equation}
The best-in-basis badness is always smaller than the true badness, as it does not include the errors associated with the limitation of the basis used for the fitted model. The badnesses provide a more complete description of the self-calibration performance than the RMS error on the fitted sources' fluxes, as the RMS source error only applies to the bright sources within the survey selected for the self-calibration procedure.

@@

\subsection{Self-Calibration Performance}
The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets. With the Survey A dataset it does not converge in any reasonable time period, even when the system is started close to the optimum fit. The fitted instrument response solutions from the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}. In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis''). The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.

@@

With the Survey C and D datasets, the instrument response is accurately recovered. As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$. The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
